### 2023-03-28 article3.qmd
update

Two pairs of BGC-Argo floats were deployed in the Labrador Sea in late May 2022. The first pair (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second pair (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).

@@

```{r, map of 4 floats, fig.height=6, fig.width=6}
#| label: fig-float-position
#| fig-cap: Trajectory of each float since their deployment (black triangles). Dots represent ascending profiles. The profile that concludes the time series for each float is highlighted by a black contour.

@@

p +
  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
  scale_color_brewer(palette = 'Dark2') + theme_bw() +
  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA)))) +
  annotation_scale(location = "br") + 
  annotation_north_arrow(location = "tr", which_north = "true") 
```

>>>>

Two pairs of BGC-Argo floats were deployed in the Labrador Sea in late May 2022 (@fig-float-position). The first pair (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second pair (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).

@@

```{r, map of 4 floats, fig.height=5, fig.width=6}
#| label: fig-float-position
#| fig-cap: Trajectory of each float since their deployment (black triangles). Dots represent ascending profiles. The profile that concludes the time series for each float is highlighted by a black contour.

@@

p +
  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
  scale_color_brewer(palette = 'Dark2') + theme_bw() +
  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA)))) +
  annotation_scale(location = "br") + 
  annotation_north_arrow(location = "tr", which_north = "true") 
```
```{=latex}
\newpage
```

### 2023-03-28 uvp6_part.qmd
update

Building the zooplankton classification model for the UVP6 was the first task of my PhD. I knew nothing about machine learning before I started, so I spent the first three weeks reading a machine learning book to get familiar with the concepts. To be even more honest, I also knew nothing about zooplankton either, so there was something else to learn.

In machine learning, the "no free lunch theorem" [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is]. In other words, it means that you need to get to know your data and your hands dirty to find a model that works. The first step consisted in developing a full data processing pipeline for UVP5 images. The aim was to be prepared for the UVP6 image data set (which was not ready at that time) to train the true, final, classification model. To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labeled particle/zooplankton images and some hyperparameters (e.g. learning rate). It made my life easier when I got the final UVP6 data set and should serve other future users in the same way.

Classifying zooplankton on UVP images is difficult for human operators so building a model to automatically classify them is already a challenge, but adding technical and energy constraints to it was the icing on the cake. Below is the first draft of a publication on the classification model.

@@

Millions of images of oceanic particles have been acquired with ship-based imaging instruments. The identification of objects on images is time consuming for human operators but can be accelerated using artificial intelligence algorithms. Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but have limited battery life and must spend as little time as possible at the ocean surface to transmit their data, which means that images cannot be transmitted. In this paper, we describe an image classification model embedded in the Underwater Vision Profiler 6 (UVP6) that is appropriate to be run on the limited power and bandwidth of AUVs. To this end, we computed simple global features of images of objects of interest and gradient boosted trees as a classifier. We compared its performance with a state-of-the-art convolutional neural network, which is much more energy demanding. In addition, we show how we dealt with class imbalance (\> 80% of images are marine snow) and what were the compromises to improve the classification scores of biological classes. Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyze preliminary results from BGC-Argo floats deployed in the Labrador Sea.

@@

It is quite difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g. resolution), the number of classes (trade off between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g. number of images per class, diversity, balance), the scientific objective (e.g. precision-recall trade off), the type and number of features and the algorithm used. In many cases, RF seemed to be the best classifier [@Gorsky2010-ca; @Faillettaz2016-wr] but in embedded applications, energy constraints need to be taken into account. Therefore, the model needs to be small and the classification should be easy and fast. Gradient Tree Boosting (GTB) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity). For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GTB, also known to be one of the best assets to win Kaggle's machine learning competitions.

@@

![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}
```{=latex}
\newpage
```

#### Data preparation

@@

```{=latex}
\newpage
```

#### How to use the CR in real applications?

@@

```{=latex}
\newpage
```

#### Prospects

>>>>

Building the zooplankton classification model for the UVP6 was the first task of my PhD. To be honest, I knew nothing about machine learning before I started this so I spent the first three weeks reading a nice machine learning book in the corner of my team leader's office. To be even more honest, I basically knew nothing about zooplankton either. I think it is time to admit it.

In machine learning, the no free lunch theorem [@Wolpert1996-hw] states that if you do not make any assumption about your data, then there is no model that is a priori guaranteed to work better [@Geron2019-is]. In other words, it means that you need to get your hands dirty to find a model that works. The first step consisted in developing a full data processing pipeline using UVP5 images. The aim was to be prepared for the UVP6 image data set (not ready at that time) to train the true final classification model. To this end, I developed a python package called [UVPec](https://github.com/ecotaxa/uvpec) that is able to create a full classification model based on image folders with labeled particle/zooplankton images and some hyperparameters (e.g. learning rate). It really made my life easier when I got the final UVP6 data set and it saved me a lot of time.

In the end, I am rather satisfied with the results. Classifying zooplankton is already difficult for human operators (at least, for me) so building a model to automatically classify them is one challenge but adding technical and energy constraints to it was the ice on the cake. Below is the first draft of a publication on the classification model.

@@

Millions of images of oceanic particles have been acquired with ship-based imaging instruments. The task of labeling images is time consuming but necessary to train classification models using artificial intelligence algorithms. Autonomous Underwater Vehicles (AUVs) such as Biogeochemical-Argo floats (BGC-Argo) have revolutionized our understanding of oceanic systems but they are flawed by their battery life and by the time they can spend at the ocean surface to transmit their data. In this paper, we show how we developed the first zooplankton classification model for AUVs using the Underwater Vision Profiler 6 (UVP6). To this end, we used a features-based classifier called eXtreme Gradient Boosting (XGBoost) adapted to specific energy constraints imposed by profiling floats and we compared its performance with a state-of-the-art convolutional neural network as if energy was not a limiting factor. In addition, we show how we dealt with a highly imbalanced data set towards marine snow (\> 80% of detritus) and what were the compromises to improve the classification scores of biological classes. Finally, we show the classification report (precision and recall for each label) of the final classification model and use it to analyse some preliminary results from BGC-Argo floats deployed in the Labrador Sea.

@@

It is quite difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g. resolution), the number of classes (trade off between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g. number of images per class, diversity, balance), the scientific objective (e.g. precision-recall trade off), the type and number of features and the algorithm used. In many cases, RF seemed to be the best classifier [@Gorsky2010-ca; @Faillettaz2016-wr] but in embedded applications, energy constraints need to be taken into account. Therefore, the model needs to be small and the classification should be easy and fast. Gradient Boosting Tree (GBT) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity). For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GBT, also known to be one of the best assets to win Kaggle's machine learning competitions.

@@

![UVP5 and UVP6 profile coverage used in this study. UVP5 profiles used in the test set are depicted by green crosses.](image/uvp6/map_of_profiles.png){#fig-map-of-profiles}

```{=latex}
\newpage
```
#### Data preparation

@@

```{=latex}
\newpage
```
#### How to use the CR in real applications?

@@

```{=latex}
\newpage
```
#### Prospects
